---
title: "Spatial Regression"
author: "Brandon Gulley"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}

library(biscale)
library(car)
library(cleangeo)
library(cowplot)
library(geosphere)
library(maps)
library(mapproj)
library(maptools)
library(spatialreg)
library(spdep)
library(tidyverse)
library(visreg)

```

install.packages("maptools", repos="http://R-Forge.R-project.org")

``` {r data, echo=FALSE, include=TRUE}

data <- read.csv('https://raw.githubusercontent.com/chrismgentry/Spatial-Regression/master/Data/childpov18_southfull.csv', 
                 colClasses = c("character", "character", "character", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric"))

```

``` {r rename, echo=FALSE, include=TRUE}

names(data)[names(data)=="X2016.child.poverty"] <- "child.pov.2016"

```

``` {r AR subset, echo=FALSE, include=FALSE}

ar_pov <- data %>% subset(State == "AR")

summary(ar_pov)

equation <- child.pov.2016 ~ rural + urban + lnmanufacturing + lnag + 
  lnretail + lnhealthss + lnconstruction + lnlesshs + 
  lnunemployment + lnsinglemom + lnblack + lnhispanic + 
  lnuninsured + lnincome_ratio + lnteenbirth + lnunmarried

options(scipen = 5)

```

# **Ordinary Least Squares Model**

``` {r model ols, echo=FALSE, include=TRUE}

ols <- lm(equation, data=ar_pov)

summary(ols)

par(mfrow=c(4,4), mar=c(2, 2, 2, 2))
visreg(ols)

```

# **Contiguity**

``` {r contiguity, echo=FALSE, include=TRUE}

fips <- county.fips
fips.codes <- separate(data = fips, col = polyname, into = c("state", "county"), sep = ",")
ar_fips <- subset(fips.codes, state=="arkansas", select=fips)

arkansas <- maps::map(database = "county", regions = "arkansas", fill=T, plot=F)
ar_sp = map2SpatialPolygons(arkansas,ar_fips$fips,CRS("+proj=longlat"))

cleaned <- clgeo_Clean(ar_sp)
neighb.data <- poly2nb(cleaned, queen=T)
cont.neighb <- nb2listw(neighb.data,style="W", zero.policy = TRUE)

```



``` {r moran, echo=FALSE, include=TRUE}

lm.morantest(ols, cont.neighb)

lm.LMtests(ols, cont.neighb, test="all")

SLX.model <- spatialreg::lmSLX(equation, data=ar_pov, cont.neighb)

summary(SLX.model)

summary(spatialreg::impacts(SLX.model, cont.neighb), zstats = TRUE)[["pzmat"]]

```

**STOP! Neither LM-Err or LM-Lag are significant, so we are keeping the OLS results! Looking at the OLS results, it seems that being born to a teenage mother is the biggest predictor of child poverty (who would've guessed). Just for funsies though, I still ran the other models.**

# **Spatial Lag Model**

``` {r lag, echo=FALSE, include=TRUE}

sp.lag.model <- spatialreg::lagsarlm(equation, data=ar_pov, cont.neighb)

summary(sp.lag.model, Nagelkerke = TRUE)

summary(spatialreg::impacts(sp.lag.model, listw = cont.neighb, R=100), zstats = TRUE)[["pzmat"]]

```

# **Spatial Error Model**

``` {r spatialerror, echo=FALSE, include=TRUE}

sp.err.model <- spatialreg::errorsarlm(equation, data=ar_pov, cont.neighb)

summary(sp.err.model, Nagelkerke = TRUE)

spatialreg::Hausman.test(sp.err.model)

sd.err <- spatialreg::errorsarlm(equation, ar_pov, cont.neighb, etype = "emixed")
sdm <- spatialreg::lagsarlm(equation, ar_pov, cont.neighb, type = "mixed")

summary(sd.err, Nagelkerke = TRUE)

summary(spatialreg::impacts(sd.err, listw = cont.neighb, R = 100), zstats = TRUE)[["pzmat"]]

LR.Sarlm(sd.err,sp.err.model)

```

# **Building Neighbors**

``` {r neighbors, echo=FALSE, include=TRUE}

all.xy <-centroid(ar_sp)
colnames(all.xy) <- c("x","y")

#Create neighbors
all.dist.k1 <- knn2nb(knearneigh(all.xy, k=1, longlat = TRUE))
all.dist.k3 <- knn2nb(knearneigh(all.xy, k=3, longlat = TRUE))
all.dist.k5 <- knn2nb(knearneigh(all.xy, k=5, longlat = TRUE))

#Determine max k distance value to neighbor
all.max.k1 <- max(unlist(nbdists(all.dist.k1, all.xy, longlat=TRUE)))
all.max.k3 <- max(unlist(nbdists(all.dist.k3, all.xy, longlat=TRUE)))
all.max.k5 <- max(unlist(nbdists(all.dist.k5, all.xy, longlat=TRUE)))

#Calculate neighbors based on distance
all.sp.dist.k1 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k1, longlat = TRUE)
all.sp.dist.k3 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k3, longlat = TRUE)
all.sp.dist.k5 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k5, longlat = TRUE)

#Create neighbor list
all.dist.neighb.k1 <- nb2listw(all.sp.dist.k1,style="W", zero.policy = TRUE)
all.dist.neighb.k3 <- nb2listw(all.sp.dist.k3,style="W", zero.policy = TRUE)
all.dist.neighb.k5 <- nb2listw(all.sp.dist.k5,style="W", zero.policy = TRUE)

all.dist.lag.k1 <- spatialreg::lagsarlm(equation, data = ar_pov, listw = all.dist.neighb.k1)
all.dist.lag.k3 <- spatialreg::lagsarlm(equation, data = ar_pov, listw = all.dist.neighb.k3)
all.dist.lag.k5 <- spatialreg::lagsarlm(equation, data = ar_pov, listw = all.dist.neighb.k5)

summary(all.dist.lag.k1, Nagelkerke = TRUE)

```

# **Distance Error Model**

``` {r distance error, echo=FALSE, include=TRUE}

all.dist.err.k1 <- spatialreg::errorsarlm(equation, data = ar_pov, listw = all.dist.neighb.k1)
all.dist.err.k3 <- spatialreg::errorsarlm(equation, data = ar_pov, listw = all.dist.neighb.k3)
all.dist.err.k5 <- spatialreg::errorsarlm(equation, data = ar_pov, listw = all.dist.neighb.k5)

summary(all.dist.err.k1, Nagelkerke = TRUE)

```

# **Map of our Results**

``` {r map, echo=FALSE, include=FALSE}

dist.err.data <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output <- cbind.data.frame(ar_pov$FIPS,
                                    dist.err.data$fitted.values, 
                                    dist.err.data$residual, 
                                    ar_pov$child.pov.2016, 
                                    ar_pov$lnsinglemom, 
                                    ar_pov$lnuninsured, 
                                    ar_pov$lnlesshs, 
                                    ar_pov$lnincome_ratio,
                                    ar_pov$lnteenbirth,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output) <- c("fips","fitted","resid","childpov",
                        "single_mom","uninsured","less_hs","income_ratio", "teenbirth")

ar_fortify <- fortify(ar_sp)

ar_poly <- merge(x = ar_fortify, y = dist.err.output, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data <- bi_class(ar_poly, x = childpov, y = teenbirth, 
                           dim = 3, style = "quantile")

legend <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Teenage Mothers",
                    size = 6)

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

mom_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = bivariate_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-95,-80), ylim=c(25,37)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Teenage Mother Households") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
mom_pov_map

```

``` {r final, echo=FALSE, include=TRUE}

final_map <- ggdraw() +
  draw_plot(mom_pov_map, 0, 0, 1, 1) +
  draw_plot(legend, 0.400, 0.225, 0.35, 0.35)
final_map

```